One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2015-01-13 paper/results.tex
almost done with first revision

\begin{figure}
  \includegraphics[scale=0.27]{figures/polarxy_dens050}\includegraphics[scale=0.27,clip=true,trim=2.26cm 
  0cm 0cm 
  0cm]{figures/polarxy_dens110}\includegraphics[scale=0.27,clip=true,trim=2.26cm
  0cm 0cm 0cm]{figures/polarxy_dens130} 
  \caption{Visualisation of the FARGO 2D simulation in
    Fig. \ref{fargo_modeamp}. The total  
    non-axisymmetric surface density
    $\Delta\Sigma$ is shown. \label{fargo_2d}} 
\end{figure}

@@

\subsection{Dependence on the imposed temperature profile}
{\bf We show that the growth of the $m=1$ spiral is
  associated with the imposed temperature gradient  by performing a 
  series of simulations with $q\in[0,1]$,  However, to keep the Toomre $Q$
  profiles similar, we adjust the surface density power-law index such
  that $s = (3+q)/2$.  For clarity these simulations are initialised
  with $m=1$ perturbations only.  
}

% The simulations above suggest the growth of the $m=1$ spiral mode
% can be attributed to the background torque applied on linear perturbations
% as a result of an imposed temperature gradient, as discussed in
% \S\ref{global_cons}. 
% According to Eq. \ref{baroclinic_torque}, this
% torque is proportional to $q$, the power-law index used to set the
% sound-speed profile (Eq. \ref{sound-speed}) in our simulations.

Fig. \ref{fargo_varq} compares the $m=1$ spiral amplitudes as a
function of $q$. We indeed observe slower growth with decreasing
$q$. Although the figure indicates growth for the strictly isothermal 
disc ($q=0$), we did not observe a coherent one-armed spiral upon
inspection of the $m=1$ surface density field. The growth in this case
may be associated with high-$m$ modes, which dominated the 
simulation.   
% difficult to measure  
% show result of isothermal calc?
\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_fargo_varq}   
  \caption{Evolution of the $m=1$ spiral amplitude as a function of
    the {\bf negative of the imposed temperature gradient} $q$. The maximum value of the
    $m=1$ surface density in $R\in[R_1,R_2]$ is shown. 
    \label{fargo_varq}} 
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ratemax_fargo_varq}    
  \caption{Growth rates of the $m=1$ spiral mode as a function of the
    imposed sound-speed gradient $q$ (asterisks). A linear fit is also 
    plotted (dotted line). 
    \label{fargo_varq_growth}} 
\end{figure}
%maybe extend the low q simulations to get better rates?

@@

\subsection{Properties of the $m=1$ spiral and its growth}\label{fargo_m1}
{\bf Here we analyse the $q=1$ case in Fig. \ref{fargo_varq_growth} in
  more detail.  
}
%In order to characterise the $m=1$ spiral that eventually
%dominates, here we examine a simulation initialised with 
%$M=1$ in Eq. \ref{randpert}. The one-armed spiral that emerges is more
%coherent than the above simulation, which facilitates the analysis
%below.   
Fig. \ref{2d_fargo_viz} shows a snapshot of the $m=1$ surface
density {\bf of this run}. 
By measuring the $m=1$ surface density amplitude and its
pattern speed, we obtain a co-rotation radius and growth rate 
\begin{align*}
  &R_c \simeq 4.4R_0,\\
  &\gamma\simeq 0.014\Omega_k(R_0) = 0.13\Omega_p. 
\end{align*}
This one-armed spiral can be considered as low frequency because its 
pattern speed $\Omega_p \simeq 0.1\Omega_k(R_0)\lesssim 0.3\Omega$ in
$R\in[R_{1},R_{2}]$ (where it has the largest amplitude). Thus, the spiral 
pattern appears nearly stationary. The growth rate $\gamma$ is also slow
relative to the local rotation, although the characteristic growth
time $\gamma^{-1} \simeq 10P_0$ is not very long.  
\begin{figure}
  \includegraphics[width=\linewidth]{figures/polarxy2_dens120_fargo}
  \caption{Cartesian visualisation of the $m=1$ surface density
    structure in the FARGO simulation initialised with only $m=1$
    perturbations. 
    \label{2d_fargo_viz}} 
\end{figure}

@@

%snapshot shows it's tightly wound
\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_kr120_fargo}
  \caption{Normalised radial wavenumber of the $m=1$ spiral in 
    Fig. \ref{2d_fargo_viz}.\label{fargo_wavenumber}} 
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%

Using the estimated value of $R_c$, we plot in
Fig. \ref{fargo_qbarrier} the quantity $\nu^2 - 1 + Q^{-2}$, which is
required to be positive in local theory for purely wave-like
solutions to the dispersion relation (Eq. \ref{dispersion}) when the
mode frequency is given. %Note that, with the above measured wavenumber $k$, 
%Eq. \ref{dispersion} requires $\nu^2 - 1 + Q^{-2}\sim0$ and the lower
%sign (long waves) to be taken. 
Fig. \ref{fargo_qbarrier} shows two $Q$-barriers located in
the inner disc, at $R_{Qb}=R_0$ and $R_{Qb}=1.6R_0$; the bounded region is indeed 
where the $m=1$ spiral develops. This shows that the one-armed 
spiral is trapped. Note in this region, $\nu^2 - 1 + 
Q^{-2}\simeq 0.1\ll 1$, which is necessary for consistency with 
the measured wavenumber $k$ and Eq. \ref{wavenumber}.  
%Eq. \ref{dispersion} with the lower sign taken (i.e. long waves).  
There is one outer Lindblad resonance at $R_L\simeq
7.2R_0$. Thus, acoustic waves may be launched in $R\gtrsim 7.2R_0$ by
the spiral disturbance in the inner disc \citep{lin11b}. 
%implications: effect on dynamics in outer disc

@@

\subsubsection{Angular momentum exchange with the background disc}  
% The wavenumber $k = \pi G\Sigma/c_s^2$ minimises the shifted frequency
% $\sbar$ in the dispersion relation, Eq. \ref{dispersion}.  
% Physically, then, it is not surprising to 
% find perturbations of this wavenumber in a
% self-gravitating disc where $Q\sim 1$. However, according the
% local dispersion relation, $m=1$ perturbations are
% formally stable for real $k$ and $Q>1$.   
%most of the perturbation is inside co-rotation 
%so resonant interaction unlikely for instability

{\bf We explicitly show that the growth of the $m=1$ spiral is due to
  the forced temperature gradient via the background torque described
  in \S\ref{wkb}. We integrate the statement for angular momentum 
  conservation for linear perturbations, Eq. \ref{lin_ang_mom_cons},
  assuming boundary fluxes are negligible, to obtain
}
% we invoke angular momentum conservation for linear
% perturbations, Eq. \ref{lin_ang_mom_cons}. 
% Assuming fluxes are
% negligible at the disc boundaries, we can integrate
% Eq. \ref{lin_ang_mom_cons} to give
\begin{align}\label{baroclinic_torque_int}
  \frac{d}{dt}\underbrace{\int_{R_\mathrm{min}}^{R_\mathrm{max}}\jlin
    2\pi R dR}_{\jlintot} 
  =\int_{R\mathrm{min}}^{R\mathrm{max}}T_\mathrm{BG} 2\pi R dR, 
\end{align}
where we recall $T_\mathrm{BG}$ is the torque density associated with the imposed
sound-speed profile (Eq. \ref{baroclinic_torque}). We  
compute both sides of Eq. \ref{baroclinic_torque_int} using
simulation data, and compare them in Fig. \ref{fargo_angmom_ex}. There
is a good match between the two torques, especially at early times
$t\lesssim110P_0$. The average discrepancy is $\simeq 5\%$. 
The match is less good later on, when the spiral
amplitude is no longer small ($\Delta\Sigma_1\sim 0.2$ at $t=110P_0$
and $\Delta\Sigma_1\sim 0.4$ by $t=120P_0$) and linear theory becomes
less applicable. Fig. \ref{fargo_angmom_ex} confirms that the $m=1$ spiral
wave experiences a negative torque that further reduces its angular
momentum, {\bf leading to its growth.} 
This is consistent with angular momentum component
measurements (Fig. \ref {2d_angmom}).

>>>>

\begin{figure*}
  \includegraphics[scale=0.55]{figures/polarxy_dens050}\includegraphics[scale=0.55,clip=true,trim=2.26cm 
  0cm 0cm 
  0cm]{figures/polarxy_dens110}\includegraphics[scale=0.55,clip=true,trim=2.26cm
  0cm 0cm 0cm]{figures/polarxy_dens130} 
  \caption{Visualisation of the FARGO 2D simulation in
    Fig. \ref{fargo_modeamp}. The total  
    non-axisymmetric surface density
    $\Delta\Sigma$ is shown. \label{fargo_2d}} 
\end{figure*}

@@

\subsection{Dependence on the imposed temperature profile}
{\bf We show that the growth of the $m=1$ spiral is
  associated with the imposed temperature gradient  by performing a 
  series of simulations with $q\in[0,1]$.  However, to maintain similar Toomre $Q$
  profiles, we adjust the surface density power-law index such
  that $s = (3+q)/2$.  For clarity these simulations are initialised
  with $m=1$ perturbations only.  
}

Fig. \ref{fargo_varq} compares the $m=1$ spiral amplitudes as a
function of $q$. We indeed observe slower growth with decreasing
$q$. Although the figure indicates growth for the strictly isothermal 
disc ($q=0$), we did not observe a coherent one-armed spiral upon
inspection of the $m=1$ surface density field. The growth in this case
may be associated with high-$m$ modes, which dominated the 
simulation.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_fargo_varq}   
  \caption{Evolution of the $m=1$ spiral amplitude as a function of
    the {\bf negative of the imposed temperature gradient} $q$. The maximum value of the
    $m=1$ surface density in $R\in[R_1,R_2]$ is shown. 
    \label{fargo_varq}} 
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ratemax_fargo_varq}    
  \caption{Growth rates of the $m=1$ spiral mode as a function of the
    imposed sound-speed gradient $q$ (asterisks). A linear fit is also 
    plotted (dotted line). 
    \label{fargo_varq_growth}} 
\end{figure}

@@

\subsection{Properties of the $m=1$ spiral and its growth}\label{fargo_m1}
{\bf Here we analyse the $q=1$ case in Fig. \ref{fargo_varq_growth} in
  more detail.  
} 
Fig. \ref{2d_fargo_viz} shows a snapshot of the $m=1$ surface
density {\bf of this run}. 
By measuring the $m=1$ surface density amplitude and its
pattern speed, we obtain a co-rotation radius and growth rate 
\begin{align*}
  &R_c \simeq 4.4R_0,\\
  &\gamma\simeq 0.014\Omega_k(R_0) = 0.13\Omega_p. 
\end{align*}
This one-armed spiral can be considered as low frequency because its 
pattern speed $\Omega_p \simeq 0.1\Omega_k(R_0)\lesssim 0.3\Omega$ in
$R\in[R_{1},R_{2}]$ (where it has the largest amplitude). Thus, the spiral 
pattern appears nearly stationary. The growth rate $\gamma$ is also slow
relative to the local rotation, although the characteristic growth
time $\gamma^{-1} \simeq 10P_0$ is not very long.  
\begin{figure}
  \includegraphics[width=\linewidth]{figures/polarxy2_dens120_fargo}
  \caption{Cartesian visualisation of the $m=1$ surface density
    structure in the FARGO simulation initialised with only $m=1$
    perturbations. 
    \label{2d_fargo_viz}} 
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_kr120_fargo}
  \caption{Normalised radial wavenumber of the $m=1$ spiral in 
    Fig. \ref{2d_fargo_viz}.\label{fargo_wavenumber}} 
\end{figure}

Using the estimated value of $R_c$, we plot in
Fig. \ref{fargo_qbarrier} the quantity $\nu^2 - 1 + Q^{-2}$, which is
required to be positive in local theory for purely wave-like
solutions to the dispersion relation (Eq. \ref{dispersion}) when the
mode frequency is given. 
Fig. \ref{fargo_qbarrier} shows two $Q$-barriers located in
the inner disc, at $R_{Qb}=R_0$ and $R_{Qb}=1.6R_0$; the bounded region is indeed 
where the $m=1$ spiral develops. This shows that the one-armed 
spiral is trapped. Note in this region, $\nu^2 - 1 + 
Q^{-2}\simeq 0.1\ll 1$, which is necessary for consistency with 
the measured wavenumber $k$ and Eq. \ref{wavenumber}.  
There is one outer Lindblad resonance at $R_L\simeq
7.2R_0$. Thus, acoustic waves may be launched in $R\gtrsim 7.2R_0$ by
the spiral disturbance in the inner disc \citep{lin11b}.

@@

\subsubsection{Angular momentum exchange with the background disc}  
{\bf We explicitly show that the growth of the $m=1$ spiral is due to
  the forced temperature gradient via the background torque described
  in \S\ref{wkb}. We integrate the statement for angular momentum 
  conservation for linear perturbations, Eq. \ref{lin_ang_mom_cons},
  assuming boundary fluxes are negligible, to obtain
}

\begin{align}\label{baroclinic_torque_int}
  \frac{d}{dt}\underbrace{\int_{R_\mathrm{min}}^{R_\mathrm{max}}\jlin
    2\pi R dR}_{\jlintot} 
  =\int_{R\mathrm{min}}^{R\mathrm{max}}T_\mathrm{BG} 2\pi R dR, 
\end{align}
where we recall $T_\mathrm{BG}$ is the torque density associated with the imposed
sound-speed profile (Eq. \ref{baroclinic_torque}). We  
compute both sides of Eq. \ref{baroclinic_torque_int} using
simulation data, and compare them in Fig. \ref{fargo_angmom_ex}. There
is a good match between the two torques, especially at early times
$t\lesssim110P_0$. The average discrepancy is $\simeq 5\%$. 
The match is less good later on, when the spiral
amplitude is no longer small ($\mathrm{max}\Delta\Sigma_1\sim 0.2$ at $t=110P_0$
and $\mathrm{max}\Delta\Sigma_1\sim 0.4$ by $t=120P_0$) and linear theory becomes
less applicable. Fig. \ref{fargo_angmom_ex} confirms that the $m=1$ spiral
wave experiences a negative torque that further reduces its {\bf (negative)} angular
momentum, {\bf leading to its amplitude growth.} This is consistent
with angular momentum component measurements (Fig. \ref {2d_angmom}).

### 2015-01-13 paper/results_3d.tex
almost done with first revision

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent  
$n_H=2$ scale-heights at $R=R_0$. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times512$, or about $4$ cells per
$H$. Because of the reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$.
%, which
%eventually dominates in 2D simulations.

@@

Fig. \ref{3d_ampmax} plots the evolution of the $m=1$ spiral amplitudes measured
in the ZEUS-MP and PLUTO runs. We also ran simulations
with a strictly isothermal equation of state ($q=0$), which display no
growth compared to that with a temperature gradient.  This confirms
the temperature gradient effect is the same in 3D. 
%why no high m
                                %modes as in 2D for iso disc?
                                %resolution?

In the ZEUS-MP run, we observed high-$m$ disturbances developed near 
the inner boundary initially, which is likely responsible for the
growth seen at $t<50P_0$. This is a numerical artifact and effectively
seeds the simulation with a larger perturbation. Results
from ZEUS-MP are therefore off-set from PLUTO by $\sim50P_0$. However,
once the coherent $m=1$ spiral begins to grow ($t\gtrsim 100P_0$), 
we measure similar growth rates in both codes: 
\begin{align*}
  &\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
  &\gamma \simeq 0.0085\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are somewhat smaller than the 2D simulations. {\bf This is
  possibily} because of the lower resolutions adopted in 3D {\bf
  and/or because the effective Toomre parameter is larger in 3D which,
  from Eq. \ref{theoretical_rate1}, is stabilising.}

% and/or the small softening length
% employed in 2D. 
%Rc = 8.6R_0

@@

Visualisations of the 3D simulations are shown in   
Fig. \ref{3d_prelim} for the disc midplane and near the upper disc
boundary. The snapshots are chosen when the one-armed spirals in the two codes
have reached comparable amplitudes. %The agreement between the codes,
%as well as with the 2D simulations (Fig. \ref{2d_fargo_viz}), is
%satisfactory. 
Both codes show similar one-armed patterns at either height, and the
midplane snapshot is similar to the 2D simulation
(Fig. \ref{fargo_2d}). 
The largest spiral amplitude is found in the
self-gravitating region $R\in[R_1,R_2]$, independent of
height. However, notice the spiral pattern extends into 
the non-self-gravitating outer disc ($R>R_2$) at $z\sim 2H$, i.e.    
the disturbance becomes more global away from the midplane.  
%2d only adequate for self-gravitaing region 
\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.305,clip=true,trim=0cm 0cm 0cm 0cm]{figures/polarxy2_dens015_z0}  
      \includegraphics[scale=0.305,clip=true,trim=0.8cm 0cm 0cm
      0cm]{figures/polarxy2_dens015_zmax}  
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.305,clip=true,trim=0cm 0cm 0cm 0cm]{figures/pdiskxy_023_z0}
      \includegraphics[scale=0.305,clip=true,trim=0.8cm 0cm 0cm
      0cm]{figures/pdiskxy_023_zmax}
    }
  \end{center}
  \caption{Three-dimensional simulations using the ZEUS-MP (top) and 
    PLUTO (bottom) codes. The $m=1$ density component $\Delta\rho_1$
    at the midplane (left) and approximately 
    two scale-heights above the midplane (right) is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height  
    from the midplane. \label{3d_prelim}}   
\end{figure}

\subsection{Vertical structure}
Fig. \ref{3d_rz} shows the vertical structure of the one-armed 
spiral in the PLUTO run. The spiral is vertically confined to $z
\lesssim H$ at $R\sim R_0$ (the self-gravitating region). Thus, a 2D
disc model, representing dynamics near the disc midplane, is  
sufficient capture the instability. However, for $R>2R_0$ the spiral amplitude
increases away from the midplane. It remains  
small in our disc model ($|\Delta\rho_1| \lesssim 0.1$), but could become 
significant with a larger vertical domain. This means that 3D 
simulations are necessary to study the effect of the one-armed spiral
on the exterior, non-self-gravitating disc.

@@

Although Fig. \ref{3d_rz} appears to display significant vertical motion,
we measure the three-dimensionality parameter $\Theta < 10^{-2}$
(Eq. \ref{theta}), so vertical motions are insignificant compared to
horizontal motions. This supports a 2D approximation. On the other
hand, we find  $\mathrm{max}|v_z/c_s|\sim 0.2$ which, although
sub-sonic, is not very small.  
%may increase with height

>>>>

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent  
$n_H=2$ scale-heights at $R=R_0$. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times512$, or about $4$ cells per
$H$. Because of the reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$.

@@

Fig. \ref{3d_ampmax} plots the evolution of the $m=1$ spiral amplitudes measured
in the ZEUS-MP and PLUTO runs. We also ran simulations
with a strictly isothermal equation of state ($q=0$), which display no
growth compared to that with a temperature gradient.  This confirms
the temperature gradient effect is the same in 3D.

In the ZEUS-MP run, we observed high-$m$ disturbances developed near 
the inner boundary initially, which is likely responsible for the
growth seen at $t<50P_0$. This is a numerical artifact and effectively
seeds the simulation with a larger perturbation. Results
from ZEUS-MP are therefore off-set from PLUTO by $\sim50P_0$. However,
once the coherent $m=1$ spiral begins to grow ($t\gtrsim 100P_0$), 
we measure similar growth rates in both codes: 
\begin{align*}
  &\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
  &\gamma \simeq 0.0085\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are somewhat smaller than the 2D simulations. {\bf This is
  possibily} because of the lower resolutions adopted in 3D {\bf
  and/or because the effective Toomre parameter is larger in 3D
  \citep{mamat10} which, from Eq. \ref{theoretical_rate1}, is
  stabilising.}

%Rc = 8.6R_0

@@

Visualisations of the 3D simulations are shown in   
Fig. \ref{3d_prelim} for the disc midplane and near the upper disc
boundary. The snapshots are chosen when the one-armed spirals in the two codes
have reached comparable amplitudes. 
Both codes show similar one-armed patterns at either height, and the
midplane snapshot is similar to the 2D simulation
(Fig. \ref{fargo_2d}). 
The largest spiral amplitude is found in the
self-gravitating region $R\in[R_1,R_2]$, independent of
height. However, notice the spiral pattern extends into 
the non-self-gravitating outer disc ($R>R_2$) at $z\sim 2H$, i.e.    
the disturbance becomes more global away from the midplane.  
%2d only adequate for self-gravitaing region 
\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.305,clip=true,trim=0cm 0cm 0cm 0cm]{figures/polarxy2_dens015_z0}  
      \includegraphics[scale=0.305,clip=true,trim=0.8cm 0cm 0cm
      0cm]{figures/polarxy2_dens015_zmax}  
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.305,clip=true,trim=0cm 0cm 0cm 0cm]{figures/pdiskxy_023_z0}
      \includegraphics[scale=0.305,clip=true,trim=0.8cm 0cm 0cm
      0cm]{figures/pdiskxy_023_zmax}
    }
  \end{center}
  \caption{Three-dimensional simulations using the ZEUS-MP (top) and 
    PLUTO (bottom) codes. The $m=1$ density component $\Delta\rho_1$
    at the midplane (left) and approximately 
    two scale-heights above the midplane (right) is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height  
    from the midplane. \label{3d_prelim}}   
\end{figure}

\subsection{Vertical structure}
Fig. \ref{3d_rz} shows the vertical structure of the one-armed 
spiral in the PLUTO run. The spiral is vertically confined to $z
\lesssim H$ at $R\sim R_0$ (the self-gravitating region). Thus, a 2D
disc model, representing dynamics near the disc midplane, is  
sufficient capture the instability. However, for $R>2R_0$ the spiral amplitude
increases away from the midplane. It remains  
small in our disc model ($|\Delta\rho_1| \lesssim 0.1$), but could become 
significant with a larger vertical domain. This means that 3D 
simulations are necessary to study the effect of the one-armed spiral
on the exterior disc.

@@

Although Fig. \ref{3d_rz} appears to display significant vertical motion,
we measure the three-dimensionality parameter $\Theta < 10^{-2}$
(Eq. \ref{theta}), so vertical motions are insignificant compared to
horizontal motions. This supports a 2D approximation. On the other
hand, we find  $\mathrm{max}|v_z/c_s|\sim 0.2$ which, although
sub-sonic, is not very small.

### 2015-01-13 paper/summary.tex
regular update (revising)
almost done with first revision

{\bf \emph{Thermal relaxation.} The locally isothermal assumption 
} can be relaxed by including an energy equation with
{\bf a source term that restores the disc temperature over a
  characteristic timescale $t_\mathrm{relax}$}. Preliminary FARGO
simulations indicate a thermal relaxation timescale $t_\mathrm{relax} <
0.1\Omega_k^{-1}$ is needed for the one-armed spiral to
develop. However, this value may {\bf be model-dependent} and will be
explored in a follow-up study. {\bf For example,
a longer $t_\mathrm{relax}$ may be permitted with larger temperature
gradients.}

@@

{\bf \emph{Non-linear evolution}.} In the deeply non-linear regime, the
one-armed spiral may  
shock and deposit negative angular momentum onto 
the background disc. The spiral amplitude would saturate by gaining
positive angular momentum. However, if the temperature gradient is
maintained, it may be possible to achieve a balance between the gain
of negative angular momentum through the background torque, and the
gain of positive angular momentum through shock dissipation. We remark  
that fragmentation {\bf of a low-frequency spiral} is unlikely
because the co-rotation radius is outside the bulk of the spiral arm, in the non-self-gravitating
portion of the disc \citep{durisen08}. In order
to study these possibilities, improved numerical models are needed to
ensure total angular momentum conservation on timescales much longer
than that considered in this paper.

@@

{\bf\emph{Other applications of the background torque.}}     
{\bf The background torque is a generic feature in  
  discs for which the temperature is set externally.  It may therefore
  be relevant in other astrophysical contexts. }
One possibility is in Be star discs \citep{rivinius13}, 
for which one-armed oscillations may explain long-timescale variations
in their emission lines \citep[see e.g.][and references 
therein]{okasaki97,papaloizou06c,ogilvie08}. {\bf These studies 
invoke other mechanisms to produce \emph{neutral} one-armed
oscillations (e.g. rotational deformation of the star), but consider 
strictly isothermal discs.} It would be interesting to explore
the effect of a radial temperature gradient on the stability of these
oscillations.

>>>>

{\bf \emph{Thermal relaxation.} The locally isothermal assumption 
} can be relaxed by including an energy equation with
{\bf a source term that restores the disc temperature over a
  characteristic timescale $t_\mathrm{relax}$}. Preliminary FARGO
simulations indicate a thermal relaxation timescale $t_\mathrm{relax} <
0.1\Omega_k^{-1}$ is needed for the one-armed spiral to
develop. However, this value {\bf is likely model-dependent. For example,
a longer $t_\mathrm{relax}$ may be permitted with larger temperature
gradients. This issue, together with a parameter survey, will be
considered in a follow-up study.}

@@

{\bf \emph{Non-linear evolution}.} In the deeply non-linear regime, the
one-armed spiral may  
shock and deposit negative angular momentum onto 
the background disc. The spiral amplitude would saturate by gaining
positive angular momentum. However, if the temperature gradient is
maintained, it may be possible to achieve a balance between the gain
of negative angular momentum through the background torque, and the
gain of positive angular momentum through shock dissipation. We remark  
that fragmentation is unlikely because the co-rotation radius is
outside the bulk of the spiral arm \citep{durisen08}. In order
to study these possibilities, improved numerical models are needed to
ensure total angular momentum conservation on timescales much longer
than that considered in this paper.

@@

{\bf\emph{Other applications of the background torque.}}     
{\bf The background torque is a generic feature in  
  discs for which the temperature is set externally.  It may therefore
  be relevant in other astrophysical contexts. }
One possibility is in Be star discs \citep{rivinius13}, 
for which one-armed oscillations may explain long-timescale variations
in their emission lines \citep[see e.g.][and references 
therein]{okasaki97,papaloizou06c,ogilvie08}. {\bf These studies 
invoke alternative mechanisms to produce \emph{neutral} one-armed
oscillations (e.g. rotational deformation of the star), but consider 
strictly isothermal discs.} It would be interesting to explore
the effect of a radial temperature gradient on the stability of these
oscillations.
